---
title: "index"
author: "Alanna Kaiser"
date: "5/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About Me

My name is Alanna Kaiser and I am an MSW candidate at the University of Michigan. In U-M's social work program, I am concentrating my studies in Social Policy and Evaluation, with a particular interest in communities and social systems. I have professional experience working in education, housing, and community-based development and neighborhood revitalization.  

### Current Work

This summer, I am working as a Measurement and Evaluation intern with Habitat for Humanity Greater San Francisco. In this role, I am specifically conducting research and implementing a program evaluation to aid the Neighborhood Revitalization (NR) and Homeowner Development Departments (HDD), respectively.  With NR, I am compiling evidence to support Habitat’s work in green space improvement in the Bayview Neighborhood of San Francisco. With HDD, I am conducting an internal evaluation to map the lifecycles of each program, highlighting key stages, accompanying process documentation, and stakeholders engaged. 

## My Portfolio

In the coming weeks, I will add content to this online space that showcases my growing data visualization and analysis skills. The following is a small sampling of that work:

# Data Visualization in R {.tabset .tabset-fade} 

This is a small sampling of my work depicting census data with R Statistical software, utelizing R's census api. 

## Base R

```{r}

library(tidycensus)
library(tidyverse)

RaceWAYNE<-get_acs(geography = "county",
                      variables = c(
"B02001_002",  
"B02001_003", 
"B02001_004", 
"B02001_005", 
"B02001_006", 
"B02001_007", 
"B02001_008"),
                      state = "mi",
                      county="163",
                      year=2017)

RaceWAYNE$NAME<-'Wayne'
RaceWAYNE$estimate_pct = (RaceWAYNE$estimate / sum(RaceWAYNE$estimate))*100

library(RColorBrewer)
coul <- brewer.pal(7, "Blues")

barplot(RaceWAYNE$estimate_pct, main="Racial Groups in Wayne County, MI",
        ylab = "% of Total Population", 
        names.arg=c("White", "Black/\nAfrican \nAmerican",
                    "American\n Indian & \nAlaska Native",
                    "Asian", "Native\nHawaiian &\nPacific Islander",
                    "Other", "Two or More\n Races"),
        col=coul, las=2, cex.names = .73)
par(mar=c(18,6,2,1)+.1)

```

## Ggplot2 Paired Bar Graph

```{r}

library(tidycensus)
library(tidyverse)

RaceWASH<-get_acs(geography = "county", 
                  variables = c(
                    "B02001_002", #White 
                    "B02001_003", #Black
                    "B02001_004", #American Indian and Alasks Native
                    "B02001_005", #Asian
                    "B02001_006", #Native Hawaiian and Pacific Islander
                    "B02001_007", #Other
                    "B02001_008"),#Two or more races
                  state = "mi",
                  county="161",
                  year=2017)
RaceWAYNE<-get_acs(geography = "county",
                   variables = c(
                     "B02001_002", #White 
                     "B02001_003", #Black
                     "B02001_004", #American Indian and Alasks Native
                     "B02001_005", #Asian
                     "B02001_006", #Native Hawaiian and Pacific Islander
                     "B02001_007", #Other
                     "B02001_008"),#two or more races
                   state = "mi",
                   county="163",
                   year=2017)
RaceWASH$NAME<-'Washtenaw'
RaceWAYNE$NAME<-'Wayne'
RaceWASH$estimate_pct=(RaceWASH$estimate/sum(RaceWASH$estimate))*100
RaceWAYNE$estimate_pct = (RaceWAYNE$estimate / sum(RaceWAYNE$estimate))*100
RaceBOTH<-rbind(RaceWASH,RaceWAYNE)

vcounty<-RaceBOTH$NAME
vrace<-c("White", "Black/\nAfrican American",
         "American Indian\n and Alaska Native",
         "Asian", "Native Hawaiian\n and Pacific Islander",
         "Other", "Two or More\n Races","White", "Black/\nAfrican American",
         "American Indian\n and Alaska Native",
         "Asian", "Native Hawaiian\n and Pacific Islander",
         "Other", "Two or More\n Races")
vPercent<-RaceBOTH$estimate_pct
GroupedData<-data.frame(vcounty,vrace,vPercent)

library(ggplot2)
library(ggthemes)

ggplot(data = GroupedData, aes(x=vrace,y=vPercent,fill=vcounty))+
  geom_bar(alpha=.6,colour="black", stat="identity",position = position_dodge())+
  theme_few()+
labs(title = "Breakdown of Racial Groups",
     y="% of Total Population")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(axis.title.x = element_blank(),
        legend.position = "bottom",
        plot.title = element_text(face = "bold",hjust = ".5"))+
  theme(axis.title.y = element_text(face = "bold", size=11))+
  theme(legend.title = element_text(size=11,face="bold"))+
  scale_fill_manual(name= "County", values=c("Washtenaw"= "lightblue4",
                                             "Wayne"= "cadetblue2"))

```

## gglplot2 Normal Bar Chart
 
```{r}

library(tidycensus)
library(tidyverse)

Income<-get_acs(geography = "county", 
                  variables ="B19013_001",
                  state = "mi",
                  county=c("161","163"),
                  year=2017)
library(ggplot2)
library(ggthemes)
library(scales)

ggplot(Income,
       aes(x=NAME, y=estimate, fill=NAME))+
  geom_bar(stat="identity", color= "black", alpha=.5)+
  theme_few()+
  scale_fill_manual(values=c("Washtenaw County, Michigan"= "lightblue4",
                             "Wayne County, Michigan"= "cadetblue2"))+
  theme(legend.position = "bottom")+
  labs(title = "Median Household Income", y= "Income ($)")+
  theme(legend.position="none",
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(face = "bold",hjust = ".5", size = "15"),
        axis.text.y = element_text(face="bold"),
        axis.text.x = element_text(face="bold"))+
  scale_y_continuous(labels = dollar)
```
```{r, echo=TRUE, eval=FALSE}
#Code used for Median Income graph above (excluding data collection):
ggplot(Income,
       aes(x=NAME, y=estimate, fill=NAME))+
  geom_bar(stat="identity", color= "black", alpha=.5)+
  theme_few()+
  scale_fill_manual(values=c("Washtenaw County, Michigan"= "dlightblue4",
                             "Wayne County, Michigan"= "darkturquoise"))+
  theme(legend.position = "bottom")+
  labs(title = "Median Household Income", y= "Income ($)")+
  theme(legend.position="none",
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(face = "bold",hjust = ".5", size = "15"),
        axis.text.y = element_text(face="bold"),
        axis.text.x = element_text(face="bold"))+
  scale_y_continuous(labels = dollar)

```


## ggplot2 Density Plot


```{r}
library(tidycensus)
library(tidyverse)
library(dplyr)

HousingWASH<-get_acs(geography = "tract", 
                     variables = "B25077_001",
                     state = "mi",
                     county="161",
                     year=2017)
HousingWAYNE<-get_acs(geography = "tract",
                      variables = "B25077_001",
                      state = "mi",
                      county="163",
                      year=2017)
HousingWASH$NAME<-'Washtenaw'
HousingWAYNE$NAME<-'Wayne'
BothCounties<-rbind(HousingWASH,HousingWAYNE)
BothCountiesClean<-BothCounties %>% filter(!is.na(estimate))

options(scipen = 999)

library(ggplot2)
library(ggthemes)
library(sitools)

ggplot(BothCountiesClean,
       aes(x=estimate, fill = NAME)) + geom_density(alpha=.5)+
  theme_few()+
  labs(title = "A Tale of Two Counties",
       subtitle = "Home Values, Compared",
       x="Median Home Value ($)")+
  theme(plot.title = element_text(face="bold"))+
  theme(axis.title.x = element_text(size=11,face="bold"))+
  scale_x_log10("Median Home Value ($)", labels=f2si)+
  theme(axis.title.y = element_text(size=11, face="bold"))+
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "bottom")+
  theme(plot.title = element_text(hjust = .5))+
  theme(plot.subtitle = element_text(hjust = .5))+
theme(legend.title = element_text(size=11,face="bold"))+
  scale_fill_manual(name= "County", values=c("Washtenaw"= "lightblue4",
                             "Wayne"= "cadetblue2"))
```


