---
title: "Playing with Mapping"
author: "Alanna Kaiser"
date: "6/6/2019"
output: 
  html_document:
    number_sections: no
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message= FALSE,
                      warning= FALSE)
```


##A Basic Map of Median Houshold Income in San Francisco

```{r, echo=FALSE, include=FALSE}
library(tidycensus)
library(tidyverse)

options(tigris_use_cache= TRUE)

SFIncomeData <- get_acs(state = "CA",
                        county = "San Francisco",
                        geography = "tract",
                        variables= "B19013_001",
                        geometry= TRUE
                        )

SFIncomeData %>% drop_na()

SFIncomeData = filter(SFIncomeData, !(NAME %in% c("Census Tract 9804.01, San Francisco County, California", "Census Tract 179.02, San Francisco County, California", "Census Tract 9806, San Francisco County, California")))

```

```{r}

SFIncomeData %>%
  ggplot(aes(fill=estimate))+
  geom_sf()+
  labs(title = "Median Household Income in SF by Census Tract", subtitle = "Data pulled with Census API", fill= "Median Family Income")+
   scale_fill_viridis_c(option = "magma")

```


## Interactive Income Map Made with R
Hover the cursor over this map to look at median family income in different census tracts across San Francisco. Click a census tract to see other information associated with that tract, including the Margin of Error for MFI. 

```{r}
library(mapview)

mapview(SFIncomeData, zcol = "estimate", legend = TRUE, layer.name="Median Family Income")
 
```

## Map of Bayview, SF Parks
The map below depicts all SF parks in the Bayview neighborhood, a key area of operation for Habtiat for Humanity Greater San Francisco's Neighborhood Revitalization Parks and Green Space program.

```{r}
library(png)
library(grid)

img2<-readPNG("~/Box Sync/alannakz/Git/DataViz2019/Playing with Mapping/BayviewParks.png")
grid.raster(img2)

```

